Reproduction frequency and offspring survival decline in elephant seals past prime age

### 2024-07-12 paper/paperv3.qmd
more edits from marm

ggsave("outputs/repro_senesce.svg", width = 11, height = 6, units = "in")
```

@@

seal_phenology <- sealdat %>% 
  mutate(BA = firstobsbreeddoy, 
         BD = firstobsbreeddoy + breeddur, 
         MA = lastobsmoltdoy - moltdur, 
         MD = lastobsmoltdoy) %>% 
  drop_na(age, BA, BD, MA, MD)

@@

h1_aic <- do.call(AIC, h1_selective_models) %>% 
  as_tibble() %>% 
  mutate(model = c("Base", 
                   "Selective appearance", 
                   "Selective disappearance", 
                   "Selective appearance and disappearance"),
         dAIC = AIC - min(AIC),
         rellik = exp(-0.5 * dAIC),
         weight = rellik / sum(rellik)) %>% 
  relocate(model)

# Selective appearance model, first repro 95% ci
h1_selapp_coef <- summary(h1_selective_models[[2]])$coef["firstrepro", ]
make_95 <- function(ci95) {
  sprintf("95%% CI: %0.3f - %0.3f", ci95[1], ci95[2])
}
h1_firstrepro_95ci <- (h1_selapp_coef["Estimate"] + c(-1.96, 1.96) * h1_selapp_coef["Std. Error"]) %>% 
  make_95()

h1_aic
```

@@

missing_mod <- glmer(unobserved ~ age10 + (1 | year) + (1 | animalID), 
                     missing_seals, 
                     family = binomial)
missing_pred <- ggpredict(missing_mod,
                          terms = "age10 [all]") %>% 
  as_tibble() %>% 
  mutate(age = x * 10 + age_senesce,
         unobserved = predicted)

missing_age_coef <- sprintf(
  "%0.3f \u00B1 %0.3f",
  summary(missing_mod)$coef["age10", "Estimate"],
  summary(missing_mod)$coef["age10", "Std. Error"] * 1.96
)

@@

missing_seals %>% 
  group_by(age) %>% 
  summarize(unobserved_mean = mean(unobserved),
            unobserved_se = sqrt(unobserved_mean * (1 - unobserved_mean) / n()),
            unobserved_lwr = unobserved_mean - unobserved_se,
            unobserved_upr = unobserved_mean + unobserved_se) %>% 
  ggplot(aes(age)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              missing_pred,
              alpha = 0.2, 
              fill = "#de77ae") +
  geom_line(aes(y = unobserved), 
            missing_pred, 
            color = "#de77ae",
            linewidth = 2) +
  geom_pointrange(aes(y = unobserved_mean, 
                      ymin = unobserved_lwr, ymax = unobserved_upr)) +
  scale_x_continuous("Age", breaks = seq(5, 19, by = 4)) +
  scale_y_continuous("Unobserved proportion", 
                     labels = scales::label_percent(accuracy = 1))
```

>>>>

ggsave("outputs/repro_senesce.svg", width = 13, height = 6, units = "in")
```

@@

seal_phenology <- sealdat %>% 
  filter(year >= 2011) %>% 
  mutate(BA = firstobsbreeddoy, 
         BD = firstobsbreeddoy + breeddur, 
         MA = lastobsmoltdoy - moltdur, 
         MD = lastobsmoltdoy) %>% 
  drop_na(age, BA, BD, MA, MD)

@@

get_old_coef <- function(model) {
  old_coef <- summary(model)$coefficients["age10:age_catOld", "Estimate"]
  old_se <- summary(model)$coefficients["age10:age_catOld", "Std. Error"]
  ci <- old_coef + c(-1.96, 1.96) * old_se
  tibble(mean = old_coef, 
         lwr = ci[1], 
         upr = ci[2])
}

h1_selective_coef <- map(h1_selective_models, get_old_coef) %>% 
  list_rbind() %>% 
  mutate(model = c("Base", "Appearance", "Disappearance", "Combined"),
         hypothesis = "H1")

h2_selective_coef <- map(h2_selective_models, get_old_coef) %>% 
  list_rbind() %>% 
  mutate(model = c("Base", "Appearance", "Disappearance", "Combined"), 
         hypothesis = "H2")

selective_coef <- rbind(h1_selective_coef, h2_selective_coef) %>% 
  mutate(model = factor(model, levels = c("Base", "Appearance", "Disappearance", "Combined")))

ggplot(selective_coef, aes(model, mean, color = hypothesis)) + 
  geom_pointrange(aes(ymin = lwr, ymax = upr), position = position_dodge(width = 0.2)) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_color_manual(values = c("firebrick", "cornflowerblue")) +
  expand_limits(y = 0) + 
  labs(y = "Coefficient", x = "Model") + 
  theme(legend.position = "none")

@@

missing_mod <- glmer(unobserved ~ age10 + (1 | year_fct) + (1 | animalID), 
                     missing_seals, 
                     family = binomial)

summary(missing_mod)

@@

missing_years <- count(missing_seals, year_fct)
missing_pred <- expand_grid(age = 5:19, 
                            year_fct = factor(1989:2022)) %>% 
  mutate(age10 = (age - age_senesce) / 10) %>% 
  mutate(predicted = predict(missing_mod, 
                             newdata = missing_grid,
                             re.form = ~ (1 | year_fct)),
         se = predict(missing_mod, 
                      newdata = missing_grid,
                      se.fit = TRUE,
                      re.form = NA)$se.fit,
         conf.low = predicted - 1.96 * se,
         conf.high = predicted + 1.96 * se) %>% 
  left_join(missing_years, by = "year_fct") %>% 
  group_by(age) %>% 
  summarize(across(c(predicted, conf.low, conf.high), mean)) %>% 
  mutate(across(c(predicted, conf.low, conf.high), family(missing_mod)$linkinv))

missing_pred_years <- expand_grid(age = 5:19, 
                            year_fct = factor(1989:2022)) %>% 
  mutate(age10 = (age - age_senesce) / 10) %>% 
  mutate(predicted = predict(missing_mod, 
                             newdata = missing_grid,
                             type = "response",
                             re.form = ~ (1 | year_fct)))

missing_pred

missing_seals %>% 
  group_by(age) %>% 
  summarize(unobserved_mean = mean(unobserved),
            unobserved_se = sqrt(unobserved_mean * (1 - unobserved_mean) / n()),
            unobserved_lwr = unobserved_mean - unobserved_se,
            unobserved_upr = unobserved_mean + unobserved_se) %>% 
  ggplot(aes(age)) +
  geom_line(aes(y = predicted, group = year_fct), 
            missing_pred_years, 
            alpha = 0.2) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high),
              missing_pred,
              alpha = 0.2,
              fill = "#de77ae") +
  geom_line(aes(y = predicted),
            missing_pred,
            color = "#de77ae",
            linewidth = 2) +
  geom_pointrange(aes(y = unobserved_mean, 
                      ymin = unobserved_lwr, ymax = unobserved_upr)) +
  scale_x_continuous("Age", breaks = seq(5, 19, by = 4)) +
  scale_y_continuous("Unobserved proportion", 
                     labels = scales::label_percent(accuracy = 1),
                     limits = c(0, NA))
```
